\section{Functional Limits and continuity}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Functional Limits}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Exercises}

%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{4.2.1}
\\

\begin{enumerate}[(a)]

\item Supply the details for how Corollary 4.2.4 part (ii) follows from the Sequential Criterion for Functional
Limits in Theorem 4.2.3 and the Algebraic Limit Theorem for sequences proved in Chapter 2.
\\

For reference, the Algebraic Limit Theorem is Theorem 2.3.3.

The Sequential Criterion for Functional Limits here tells us that since we know that since
$\lim_{x\rightarrow c} f(x_n) \rightarrow L$ and
$\lim_{x\rightarrow c} g(x_n) \rightarrow M$,
then we know that $(x_n) \rightarrow c$ ($x_n \neq c$ for all $n$)
and that $f(x_n) \rightarrow L$ and $g(x_n) \rightarrow M$.

Since the later two expressions are sequences on their own, we can use now the Algebraic Limit Theorem,
$\lim \left(a_n + b_n\right) = a+b$, implying that
$\lim_{x\rightarrow c} \left( f(x_n) + f(x_n)\right) \rightarrow L + M$.


\item Now, write another proof of Corollary 4.2.4 part (ii) directly from Definition 4.2.1 without using the
Sequential Criterion for Functional Limits in Theorem 4.2.3.

To do the same proof without the Sequential Criterion for Functional Limits we ought to go back to $\epsilon-\delta$ proofs.

Since we know that $\lim_{x\rightarrow c} f(x_n) \rightarrow L$ and
$\lim_{x\rightarrow c} g(x_n) \rightarrow M$, then we can define a $\delta_f$ such that when $0 < |x-c| < \delta_f$ $|f(x) - L|<\epsilon_f$
(for any $\epsilon_f > 0$)
and a $\delta_g$ sucht that when $0 < |x-c| < \delta_g$ $|g(x) - M| < \epsilon_g$ (for any $\epsilon_g > 0$).

Then if we look for a $\delta = \min\left\{\delta_f, \delta_g \right\}$ such that
$$
\left| \left(f(x)+g(x)\right) - \left(L+M\right) \right| \leq
\left|f(x) - L \right| + \left| g(x) - M \right| < \epsilon_f + \epsilon_g = \epsilon
$$

If you notice, this is pretty much the same sort of logic we used to prove the Algebraic Limit Theorem, so we could
have a guess for what the proof for the next bit could be.

\item Repeat (a) and (b) for Corollary 4.2.4 part (iii).

The proof using the Sequential Criterion for Functional Limits is essentially the same as in part (a).
(You should definetely write it out or read it out loud though!)
So let's carry on with the proof without it.

Similar to what we did while proving the Algebraic Limit Theorem, let's work backwards from what we want.

\begin{align*}
\left| f(x)g(x) - LM \right| &=  \left| f(x)g(x) - Lg(x) + Lg(x) - LM \right| \\
&\leq \left| f(x)g(x) - Lg(x) \right| + \left| Lg(x) - LM \right| \\
&= |g(x)| \left| f(x) - L \right| + |L|\left| g(x) - M \right|
\end{align*}

Again, we want to find the "worst case scenario" for $g(x)$.
And since $g(x_n)$ is a convergent sequence, then it will have some bound $M$.
If we also recall our $\delta_f$ and $\delta_g$ from part (b), then we have

\begin{align*}
\left| f(x)g(x) - LM \right| &\leq |g(x)| \left| f(x) - L \right| + |L|\left| g(x) - M \right| \\
&\leq M \left| f(x) - L \right| + |L|\left| g(x) - M \right| \\
&< |M| \epsilon_f + |L| \epsilon_g \\
&= |M| \left( \frac{\epsilon}{2|M|} \right) + |L|\left( \frac{\epsilon}{2|L|} \right) = \epsilon
\end{align*}

At the end there, we went ahead and figured out how to define $\epsilon_f$ and $\epsilon_g$ in terms of some arbitrary $\epsilon$
that could be used for the entire thing.

\end{enumerate}



%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{4.2.2}
\\

For each stated limit, find the largest possible $\delta$-neighborhood that is a proper response to the $\epsilon$-challenge.

\begin{enumerate}[(a)]
\item $\lim_{x\rightarrow 3} \left( 5x - 6 \right) = 9$, where $\epsilon = 1$.

We are looking for a $\delta$ such that when $0 < |x-3| < \delta$, $|5x -6 -9| = |5x- 15|= |5||x-3| = 5|x-3| < 1$.

If we choose $\delta = 1/5$, then $5|x-3| < 5\delta = 1 = \epsilon$.

\item $\lim_{x\rightarrow 4} \sqrt{x} = 2$, where $\epsilon = 1$.

We are looking for a $\delta$ such that when $0 < |x-4| < \delta$, $|\sqrt{x}-2| < \epsilon$.

One way to try this one out would be not note that
$$
|\sqrt{x}-2| = \frac{|x-4|}{|\sqrt{x}+2|} \leq \frac{|x-4|}{2} < \frac{\delta}{2}
$$

So we could choose $\delta = 2$ and have a proper response to our $\epsilon = 1$ challenge.
But because we used an inequality, $\delta = 2$ is not the largest possible $\delta$-neighborhood available.

After doing some internet research, we saw no better answer than to proceed with some numerical experiments.
We are looking for a $\delta$ that will put us plus or minus 1 from $y = \sqrt{x} = 2$, so $1 < y = \sqrt{x} < 3$
or $1 < x < 9$.
This last expression is equivalent to $4-\delta < x < 4+\delta$ or $-(\delta +4) < x < \delta + 4$.
Meaning that we have two options 3 or 5 (do the algebra to solve for $\delta$ using the last expression or just count using $1 < x < 9$).

Now, we pick the smallest of the two possible values for $\delta$ because that one will give us an entire $V_\delta (c)$ that will result
in $|\sqrt{x} - c| < \epsilon$ (imagine the graph - neighborhoods are symmetric so you ought to avoid a neighborhood that will give you $x$
values that will put you outside of the desired $\epsilon$ tolerance).

Numerically, if we chosed the largest value for delta, $\delta = 5$, then $0 < |x-4| < 5$ implies $-5 < x-4 < 5$ which means that $x$ could have any value
$-1 < x < 9$, and $-1$ is not a valid value in this case.

This analysis was based on
\href{https://math.libretexts.org/Bookshelves/Calculus/Calculus_3e_(Apex)/01%3A_Limits/1.02%3A_Epsilon-Delta_Definition_of_a_Limit}{Calculus 3e Apex: Epsilon-Delta Definition of a Limit}


\item $\lim_{x\rightarrow \pi} [[x]] = 3$, where $\epsilon = 1$.
(The function $[[x]]$ returns the greatest integer less than or equal to $x$.)

Once again, we are looking for a $\delta$ such that when $0 < |x- \pi| < \delta$, $|[[x]] - 3| < \epsilon = 1$.

Doing a bit of thinking, any $x \in (2,5)$ will result in $[[x]] = 2, 3, 4$.
All of these values are within a distance of 1 from 3.
If we tried then $|2-\pi|$ or $|5-\pi|$ however, we would have something like $|[[2.001]]-3|$ or $|[[4.999]]-3|$
which is are not strictly smaller than 1.

Because of the nature of $[[x]]$ (you should draw it), the way to narrow our domain and stay some distance less than one unit
from $3$ is to have $x$ such that $x\in (3,4)$.
Specifically, the largest $\delta$-neighborhood would be $\pi - 3$, since any value greater than that would put $x$ in the domain of
$x\in (2,3)$ which would make $[[x]]=2$.

\item $\lim_{x\rightarrow \pi} [[x]]$, where $\epsilon = 0.01$.

Same as above: $\delta = \pi - 3$.

\item $\lim_{x\rightarrow 0} x^3 = 0$

A $\delta$ such that when $0 < |x-0| = |x| < \delta$ $|x^3 - 0| = |x|^3 < \epsilon$ for every $\epsilon > 0$
means that we can choose $\delta = \epsilon^{1/3}$.

\item $\lim_{x\rightarrow 2} x^3 = 8$

A $\delta$ such that when $0 < |x-2| < \delta$ $|x^3 - 8| < \epsilon$ for every $\epsilon > 0$.

Following Example 4.2.2 (ii), we can agree that $|x-2|<1$, so $x=1,3$.
Then we can factor $x^3 -8$ into $(x-2)(x^2 + 2x + 4)$ if we try $x=1$ then $x^2 + 2x + 4 = 7$.
If we try $x=3$, $x^2 + 2x + 4 = 19$.
Out of these two options, choosing 7 would lead us to chose $\delta = \epsilon /7$, which is a better upper bound than $\epsilon / 19$.
So $\delta = \min(1, \epsilon/7)$ leads us to $|x-2| |x^2 +2x + 4| = \frac{\epsilon}{7} \cdot 7 = \epsilon$

\end{enumerate}



%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{4.2.3}
\\

Review the defintion of Thomae's function $t(x)$ from Section 4.1.

The Thomae function is defined as
$$
t(x) =
\begin{cases} 
    1   & x = 0 \\
    1/n & x=m/n \in \mathbb{Q} \setminus \{0\} \text{ is in lowest terms with } n>0 \\
    0   & x \not\in \mathbb{Q} 
\end{cases}
$$

\begin{enumerate}[(a)]
\item Construct three different sequences $(x_n)$, $(y_n)$, and $(z_n)$, each of which converges to 1 without using 1 as a term in
the sequence.

$(x_n) = n/(n+1)$, $(y_n) = (n+1)/n$, and $(z_n) = 1 + 1/n^2$

\item Now, compute $\lim t(x_n)$, $\lim t(y_n)$, and $\lim t(z_n)$.

Writing out a couple terms for each: if
$(x_n) = \left\{ \frac{1}{2}, \frac{2}{3}, \frac{4}{3}, \frac{4}{5}, \ldots , \frac{99}{100}, \frac{100}{101}, \ldots \right\}$, then
$t(x_n) = \left\{ 1, \frac{1}{2}, \frac{1}{3}, \frac{1}{4}, \ldots , \frac{1}{99}, \frac{1}{100}, \ldots \right\}$.

If $(y_n) = \left\{\frac{2}{1}, \frac{3}{2}, \frac{4}{3}, \frac{5}{4}, \frac{6}{5}, \ldots , \frac{10}{9}, \frac{11}{10}, \ldots \frac{100}{99}, \frac{101}{100}, \ldots \right\}$, then
$t(y_n) = \left\{ 1, \frac{1}{2}, \frac{1}{3}, \frac{1}{4}, \ldots , \frac{1}{99}, \frac{1}{100}, \ldots \right\} = t(x_n)$.

If $(z_n) = \left\{1+\frac{1}{1}, 1+\frac{1}{4}, 1+\frac{1}{9}, \ldots , 1+\frac{1}{100}, \ldots \right\}$, then
$t(z_n) = t(y_n) = t(z_n)$.

Thus we have $\lim t(x_n) = \lim t(y_n) = \lim t(z_n) = 0$.

\item Make an educated conjecture for $\lim_{x\rightarrow 1} t(x)$, and use Definition 4.2.1B to verify the claim.
(Given $\epsilon > 0$, consider the set of points $\{ x\in\mathbb{R} : t(x) \geq \epsilon \}$.
Argue that all the points in this set are isolated.)

Our claim will be that $L=0$ when $x\rightarrow 1$.
Which means that for every $V_{\epsilon} (L)$ there is a $V_{\delta} (1)$ such that for all $x \in V_{\delta} (1)$, $f(x) \in V_{\epsilon} (L)$.
Remeber that this is equivalent to saying that for every $\epsilon >0$, there exist a $\delta$ such that when $0 < |x-1| < \delta$,
$|t(x)| < \epsilon$.

Remeber, that we need to find a $\delta$ for the $\epsilon$ challenge.
First, we consider some arbitrary $\epsilon > 0$ and some $x in \mathbb{Q}$.
So $x = m/n$ and $t(x) = 1/n$.
Then $0 < |x-1| = |\frac{m}{n} - \frac{n}{n}| = |\frac{m-n}{n}| < \delta$ and that entails that $1/n < \epsilon$.
Though notice that $1/n \leq |x-1| = |\frac{m-n}{n}| < \delta$.
So we can set $\delta = \epsilon$.

\end{enumerate}



%%%%%%%%%%%%%%%%%%%%%
\textbf{4.2.4}
\\

Consider the reasonable but erroneous claim that
$$
\lim_{x\rightarrow 10} 1/[[x]] = 1/10.
$$

\begin{enumerate}[(a)]
\item Find the largest $\delta$ that represents a proper response to the challenge $\epsilon = 1/2$.

One way to go about this is to plug in numbers (or graph the whole thing) and see how things behave.
For $\epsilon = 1/2$, we can get an $f(x) \in V_{\epsilon=1/2} (L=\frac{1}{10})$ for any $x \in (-\infty, -2) \cup [2, \infty)$.

Remember that $c$ is a limit point of A if every $\delta$-neighborhood
$V_{\delta} (x=c)$ intersects A at some point other than $x$.
Since we are looking at $c=10$, the $V_\delta (c)$ neighborhood in question would be $10 \pm 8$ ($10 - 8 = 2$) since only values in
$[2,\infty)$ meet the $\epsilon$ challenge.

\item Find the largest $\delta$ that represents a proper response to the challenge $\epsilon = 1/50$.

Again, graphing the values, the only bits of the real line where we can meet this $\epsilon = 1/50$ challenge are
$x \in [9, 13)$.

This time the $x$ value that's the fartherst from $x=10$ is 13, so $\delta = 3$.

\item Find the largest $\epsilon$ challenge for which there is no suitable $\delta$ response possible.

When $[[x]] = 10$, $|f(x=10) - c| = 0$.
So a $\delta$ that leaves $x \in [10, 11)$ makes us able to meet any $\epsilon$ challenge.
However, neighborhoods are symmetric, so if we are centered at $x=10$, the smallest non-empty neighborhood would include 9.
In which case we would not be able to meet an $\epsilon$ challenge smaller than
$|\frac{1}{[[x]]} - \frac{1}{10}| = |\frac{1}{9} - \frac{1}{10}| = 1 / 90$.

\end{enumerate}



%%%%%%%%%%%%%%%%%%%%%%%
\textbf{4.2.5}
\\

Use Definition 4.2.1 to supply a proper proof for the following limit statements.

\begin{enumerate}[(a)]
\item $\lim_{x\rightarrow 2} \left(3x + 4\right) = 10$.

We want an $\epsilon > 0$ such that when
$0 < |x-2| < \delta$,
$|3x +4 - 10| = |3x -6| = 3|x-2| < \epsilon$.

We can achieve this by setting $\delta = \epsilon / 3$.

Remember that we can "control" $\delta$ but we want to put it in terms of $\epsilon$.



\item $\lim_{x\rightarrow 0} x^3 = 0$.

Now we want a $\delta$ such that when $0 < |x| < \delta$, $|x^3| = |x|^3 < \epsilon$.
If we set $\delta = \epsilon^{1/3}$ then we can meet the $\epsilon$ challenge.


\item $\lim_{x\rightarrow 2} \left(x^2 + x - 1\right) = 5$.

Now we want a $\delta$ such that when $0 < |x-2| < \delta$, $|x^2 + x -1 - 5| = |x^2 + x - 6| = |x-2||x+3| < \epsilon$.

If we define the radius of our $\delta$-neighborhood to be no more than 1, then $|x+3| \leq |3+3| = 6$ is an upper bound.
Another way to get the upper bound is to note that $|x+3| = |x-2 + 5| \sim \delta + 5 = 6$.

Now, choose $\delta = \min(1, \epsilon/6)$.

This way, if $0 < |x-2| < \delta = \min\{ 1, \epsilon/6 \}$, then
$|x-2||x+3| < 6 (\epsilon/ 6) = \epsilon$.


\textbf{Note:} if we decide that $\delta = 1$, then $|x-2| <1$ implies that $-1 < x-2 < 1$ or $1 < x < 3$.

\item $\lim_{x\rightarrow 3} 1/x = 1/3$.

We want a $\delta$ such that when $0 < |x-3| < \delta$,
$|1/x - 1/3| = \left| \frac{3-x}{3x} \right| = \frac{|x-3|}{3|x|} < \epsilon$.

If we did what we did in the previous problem, then we decide that we want $\delta = 1$ and $|x-3| < 1$
implies $-1 < x-3 < 1$ or $2 < x < 4$.
This would mean that $1/x < 1/2$ or $1/x > 1/4$, which together mean $1/4 < 1/x < 1/2$.

Thus,
$$
\frac{|x-3|}{3|x|} < \frac{1}{3} \frac{1}{2} |x-3| < \epsilon
$$

So $\delta = \min\{ 1, 6\epsilon \}$ should work.

\end{enumerate}




%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{4.2.6}
\\

Decide if the following claims are true or false, and give short justifications for each conclusion.

\begin{enumerate}[(a)]
\item If a particular $\delta$ has been constructed as asuitable response to an $\epsilon$ challenge, then any smaller positive
$\delta$ will also suffice.

This is true. As per the definition of a functional limit, if we have found a $\delta$ such that when $0 < |x-c| < \delta$
then $|f(x) - L|<\epsilon$, then any $\delta^\prime < \delta$ will also suffice as this means that
$0 < |x-c| < \delta^\prime < \delta$.


\item If $\lim_{x\rightarrow a} f(x) = L$ and $a$ happens to be in the domain of $f$, then $L = f(a)$.

Not necessarily.
A counter example: the Dirac delta function.

\item If $\lim_{x\rightarrow a} f(x) = L$, then $\lim_{x\rightarrow a} 3|f(x) - 2|^2 = 3(L-2)^2$.

Think this is true due to the algebraic limit theorem for functional limits -
limits of sums or multiplications is okay but limits of divisions
may be indeterminate, hence calling for the L'hopital (Bernoulli) rule.

\item If $\lim_{x\rightarrow a} f(x) = 0$, $\lim_{x\rightarrow a} f(x)g(x) = 0$ for any function $g$ with domain equal to the
domain of $f$.

No because $g(x)$ is rather generic so we could actually be doing a division somewhere in there and then we would need the
L'Hopital rule.

\end{enumerate}



%%%%%%%%%%%%%%%%%%%%%%%%%
\phantomsection
\label{abbott:4.2.7}
\textbf{4.2.7}
\\

Let $g : A \rightarrow \mathbb{R}$ and assume that $f$ is a bounded function in $A$
in the sense that there exists an $M > 0$ satisfying $|f(x)| \leq M$ for all $x \in A$.
Show that if $\lim_{x\rightarrow c} g(x) = 0$,
then $\lim_{x\rightarrow c} g(x)f(x) = 0$ as well.
\\

We can maybe see a little of the algebraic limit theorem here as well but remember that when looking at functions
we ought to think "functional limits" which in turn means we ought to have an $\epsilon$ and a delta $\delta$.

So think of it this way, we should be looking for a $\delta$ such that if $0 < |x-c| < \delta$,
then $|g(x)f(x)-0| = |g(x)f(x)| < \epsilon$.

We can see then that $|g(x)f(x)| \leq M |g(x)|$.

We also already know that there exists a response $\delta$ for the challenge
$|g(x)| < \epsilon_{g}$ (since $\lim_{x\rightarrow c} g(x) = 0$).
Meaning that we should be able to find a $\delta$ for the challenge $\epsilon_g = \epsilon / M$,
hence, we should be able to find a $\delta$ so that

$$
|g(x)f(x)| \leq M |g(x)| < M \frac{\epsilon}{M} = \epsilon
$$
\\~\\



%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{4.2.8}
\\

Compute each limit or state that it does not exist.
Use the tools developed in this section to justify your conclusion.

\begin{enumerate}[(a)]
\item $\lim_{x \rightarrow 2} \frac{|x-2|}{x-2}$

The limit does not exist.
There are a couple ways to go about arguing this.

The first is by remembering the defintino of $|x|$ as either $x$ or $-x$, if $x<0$.
So the above could simplify to either 1 or $-1$.

Another way is by using the sequential criterion for functional limits.
For example, if $(x_n) = 2 + 1/n$, then the limit approaches 1.
But if $(y_n) = 2 - 1/n$, then the limit is -1.

\item $\lim_{x \rightarrow 7/4} \frac{|x-2|}{x-2}$

Now, the limit does exist.
$$
\frac{|x-2|}{x-2} \Bigg|_{x=7/4} = \frac{\left|-\frac{1}{4}\right|}{-\frac{1}{4}} = -1
$$
This will be the case as long as $x<2$.
In our $\epsilon$-$\delta$ language, we want a $\delta$ such that $0 < 2-x < \delta$.

\item $\lim_{x \rightarrow 0} \left(-1\right)^{[[1/x]]}$

Since $1/x \big|_{x\rightarrow 0} = \infty$, then $[[1/x]] \big|_{x\rightarrow 0} = \infty$ as well.
But since we raising $-1$ to a power, depending on whether the exponent is even or odd, the limit will be 1 or -1, respectively.
So, again, the limit does not exist.

Using the sequential  criterion, we can look at the series $(x_n) = 1/2n$ and $(y_n) = 1/(2n+1)$, such that
$\lim x_n = \lim y_n = 0$ but $\lim f(x_n) \neq \lim f(y_n)$.


\item $\lim_{x \rightarrow 0} x^{1/3} \left(-1\right)^{[[1/x]]}$

Now, we have two competing components, $\left(-1\right)^{[[1/x]]}$, which we just saw oscillates between -1 and 1, and
$x^{1/3} \big|_{x\rightarrow 0} = 0$.
If we look back, we can use Exercise \ref{abbott:4.2.7} to conclude that this limit is 0.

Also, note that $| x^{1/3} \left(-1\right)^{[[1/x]]} | = | x^{1/3} | < \epsilon$.
Given this, any $\delta = \epsilon^3$ will be able to raise to the challenge.

\end{enumerate}



%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{4.2.9}
\\

(Infinite Limits) The statement $\lim_{x\rightarrow 0} 1/x^2 = \infty$ certainly makes intuitive sense.
To construct a rigorous definition in the challenge-response style of Definition 4.2.1 for an infinite limit statement
of this form, we replace the (arbitrarily small) $\epsilon > 0$ challenge with an (arbitrarily large) $M>0$ challenge:

Definition: $\lim_{x\rightarrow c} f(x) = \infty$ means that for all $M>0$ we can find a $\delta >0$
such that whenever $0 < |x-c| < \delta$, it follows that $f(x) > M$.

\begin{enumerate}[(a)]
\item Show $\lim_{x\rightarrow 0} 1/x^2 = \infty$ in the sense described in the previous definition.

\item Now, construct a definition for the statement $\lim_{x\rightarrow \infty} f(x) = L$. Show $\lim_{x\rightarrow \infty} 1/x =0$.

\item What would a rigorous definition for $\lim_{x\rightarrow \infty} f(x) = \infty$ look like?
Give an example of such limit.
\end{enumerate}