\section{Our Roadmap}

This roadmap is designed for the mature self-studier aiming for mathematical fluency.
It bridges the gap between computation-based learning and the abstract "rigor" required for real life.


\subsection{Phase 1: Building the Internal Map (Analysis \& Topology)}
\begin{itemize}[leftmargin=1.5em]
    \item \textbf{Understanding Analysis} -- \textit{Stephen Abbott}
    \begin{itemize}
        \item \textit{Role:} Narrative intuition. Builds the epsilon-delta "muscle memory" through clear, conversational motivation.
    \end{itemize}
    \item \textbf{Real Mathematical Analysis} -- \textit{Charles Chapman Pugh}
    \begin{itemize}
        \item \textit{Role:} The Bridge. Famous for its Chapter 2 on Metric Space Topology and its visual, "gritty" approach to rigor.
    \end{itemize}
    \item \textbf{Visual Complex Analysis} -- \textit{Tristan Needham}
    \begin{itemize}
        \item \textit{Role:} The Geometer. Essential for transforming complex variables from symbols into physical rotations and expansions.
    \end{itemize}
\end{itemize}

\subsection{Phase 2: The Operator Architecture (Linear Algebra)}
\begin{itemize}[leftmargin=1.5em]
    \item \textbf{A Student's Guide to Vectors and Tensors} -- \textit{Daniel Fleisch}
    \begin{itemize}
        \item \textit{Role:} Coordinate-free intuition. Connects physical intuition to the formal language of tensors.
    \end{itemize}
    \item \textbf{Linear Algebra Done Right} -- \textit{Sheldon Axler}
    \begin{itemize}
        \item \textit{Role:} The Rigor Check. Shifts focus from matrices to Linear Operators, the language of Functional Analysis.
    \end{itemize}
\end{itemize}

\subsection{Phase 3: The Mastery (The Rudin Trilogy)}
\begin{itemize}[leftmargin=1.5em]
    \item \textbf{Principles of Mathematical Analysis (Baby Rudin)} -- \textit{Walter Rudin}
    \begin{itemize}
        \item \textit{Role:} Polishing and Formalism. After Pugh and Abbott, this serves as the final refinement of your Real Analysis foundation.
    \end{itemize}
    \item \textbf{Real and Complex Analysis (Papa Rudin)} -- \textit{Walter Rudin}
    \begin{itemize}
        \item \textit{Role:} The Destination. Unified treatment of measure theory and holomorphic functions.
    \end{itemize}
\end{itemize}
